---
title: "Actividad 4.2. Big data"
author: "Ismael Lare David"
date: "`r Sys.Date()`"
---

# Ejercicio 1. Benchmarking

### Compare (benchmark) reading and processing the especified data with:

- duckdb (reading the file remotely)

- duckdb (reading the file locally. Include the download time in the benchmark)

- arrow (reading the file locally. Include the download time in the benchmark)

```{r}

# Instalar multiples librerias (Si no lo estan)
check.packages <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
packages<-c("openxlsx","sf","glue", "dplyr", "plyr","data.table","stringr",
            "xts","ggpubr","rstudioapi", "lubridate","DescTools","matrixStats",
            "splitstackshape", "mgsub", "textshape", "psych","fuzzyjoin", "irr", 
            "bench", "DBI", "duckdb", "arrow", "geoarrow")
check.packages(packages)


### IMPORTAR LIBRERIAS ###
library(openxlsx)
library(tidyr)
library(zoo)
library(dplyr)
library(data.table)
library(xts)
library(ggpubr)
library(purrr)
library(rstudioapi)
library(lubridate)
library(DescTools)
library(matrixStats)
library(stringr)
library(splitstackshape)
library(purrr)
library(mgsub)
library(irr)
library(janitor)
library(reshape2)
library(tibble)
library(ggplot2)
library(patchwork)
library(glue)
library(sf)
library(bench)
library(DBI)
library(duckdb)
library(arrow)
library(geoarrow)


# Select directory path
current_folder<-dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(current_folder)

#### 1) duckdb (reading the file remotely)

# Create a connection to SQL
db_conn <- dbConnect(duckdb())

# Initialize connection (necesario instalar "httpfs" si se quiere leer desde la página web)
dbExecute(db_conn, "INSTALL httpfs;")
dbExecute(db_conn, "LOAD httpfs;")


# Create query parquet file (remotely)
parquet_prep <- glue(
  "SELECT *
  FROM read_parquet('https://data-emf.creaf.cat/public/parquet/stations_data_historical/meteo_stations_2000_2024.parquet')
")

# Read parquet file remotely

remote_parquet_test <- bench::mark(dbGetQuery(db_conn, parquet_prep) %>%
  select(day,
         year,
         month,
         station_province,
         starts_with("Temperature"),
         ends_with("Humidity"),
         Precipitation,
         geom) %>% 
  filter(year == 2020) %>% 
  st_as_sf(crs = 25830)
  )


### 2) duckdb (reading the file locally. Include the download time in the benchmark)

# Create a connection to SQL
db_conn <- dbConnect(duckdb())

# Initialize connection (necesario instalar "httpfs" si se quiere leer desde la página web)
dbExecute(db_conn, "INSTALL spatial;")
dbExecute(db_conn, "LOAD spatial;")

# Create query parquet file (remotely)
locally_data <- glue(
  "SELECT *
  FROM read_parquet('DATA/meteo_stations_2000_2024.parquet')
")

locally_parquet <- bench::mark(
  DBI::dbGetQuery(db_conn, locally_data) %>% 
  select(day,
         year,
         month,
         station_province,
         starts_with("Temperature"),
         ends_with("Humidity"),
         Precipitation,
         geom) %>%
  filter( year == 2020) %>% 
  st_as_sf(crs = 25830)
)

### 3) arrow (reading the file locally. Include the download time in the benchmark)
data_parquet <- open_dataset("DATA/meteo_stations_2000_2024.parquet")

only_parquet <- bench::mark(data_parquet %>% 
  select(day,
         year,
         month,
         station_province,
         starts_with("Temperature"),
         ends_with("Humidity"),
         Precipitation,
         geom) %>%
  filter(year == 2020) %>% 
  sf::st_as_sf(crs = 25830)
)

# 4) sf with parquet (reading the file locally)
data_parquet <- open_dataset("DATA/meteo_stations_2000_2024.parquet")

data_sf_parquet <- bench::mark(data_parquet %>% 
  select(day, year, month, station_province,
        dplyr::starts_with("Temperature"),
        dplyr::ends_with("Humidity"),
        Precipitation,
        geom) %>%
  filter(year == 2020) %>% 
  st_as_sf(crs = 25830)
)

# Benchmark results
benchmark_results <- bind_rows(remote_parquet_test, locally_parquet, only_parquet, data_sf_parquet) %>%
  select(min, median, `itr/sec`,`mem_alloc`, `gc/sec`, n_itr, n_gc, total_time)

# Export results
openxlsx::write.xlsx(
  benchmark_results,
  "D:/Escritorio/Documentos/Diplomatura Ciencia de datos geográficos/R/Módulo 4/benchmark_meteo_ourense_2020.xlsx"
)

```

El método que ha presentado mayor velocidad a la hora de cargar el archivo de datos temporales ha sido la combinación de arrow para abrir el archivo parquet de forma manual junto con sf. La media de tiempo reporta unos 2.25 segundos para cargar y procesar los datos.


# Ejercicio 2.

### Write a text document (txt, doc, R markdown or Quarto) designing the steps you think would be necessary to process the same data as before, but for all municipalities in Spain, returning daily summaries of meteorological data for each municipality in 2020. Don’t write code, just write the strategy you’d follow and the steps you’d take to achieve this.

Para procesar la información meteorológica de todos los municipios de España, utilizaría una combinación de los paquetes arrow y duckdb para seleccionar y filtrar las variables que buscamos analizar. Los pasos a seguir serían los siguientes:

1 - Abrir el archivo parquet mediante la libreria "arrow" con la función "open_dataset()". Esto permitiría leer los metadatos del archivo y visualizar el contenido de las columnas.

2 - Cargar el archivo descargado mediante duck_db a través de las funciones "dbConnect()" y "dbGetQuery" y seleccionar columnas relacionadas con la fecha, municipio, las variables principales (humedad relativa, temperatura y precipitación) y geometría.

# Ejercicio 3: Parallelization I

### Using the method you prefer, repeat the previous example (coefficient bootstrapping), changing the number of cores, from 2 to the number in your system (parallel::detectCores()).
*Are always the times better as we increase the number of cores?
No, the plot show a point where the higher number of cores does not necessarily correlate with a higher speed. There is a point where the preparation of the data is heavier than the process itself.
*Can your memory hold all examples?
Yes, the process runs successfully for all the configurations.

```{r}
# libraries
library(mirai)
library(parallel)
library(foreach)
library(doParallel)
library(dplyr)

# data needed
n_repetitions <- 1e4
res_coefs <- list()
iris_data <- iris |>
  dplyr::filter(Species != "setosa")
# set the number of cores to use
doParallel::registerDoParallel(cores = 8)
# foreach loop (for illustration purposes, we check the time used)
eight_cores <- system.time(
  {res_coefs <- foreach::foreach(index = 1:n_repetitions) %dopar% {
    sample_individuals <- sample(85, 85, replace = TRUE)
    model_res <- glm(
      iris_data[sample_individuals, "Species"] ~ iris_data[sample_individuals, "Petal.Length"],
      family = binomial
    )
    coefficients(model_res)
  }}
)

test_efficiency <- rbind(four_cores, six_cores, eight_cores)

# Mclapply

# Data
iris_data <- iris |>
  filter(Species != "setosa") |>
  mutate(Species = droplevels(Species))

# Function to process one repetition
coef_function <- function(repetition) {
  
  sample_individuals <- sample(
    seq_len(nrow(iris_data)),
    nrow(iris_data),
    replace = TRUE
  )
  
  boot_data <- iris_data[sample_individuals, ]
  
  model_res <- suppressWarnings(
    glm(
      Species ~ Petal.Length,
      data = boot_data,
      family = binomial
    )
  )
  
  coefficients(model_res)
}

# Number of repetitions
n_repetitions <- 1e4

# Function to run parallel processing with a given number of cores
run_parallel <- function(n_cores) {
  
  cl <- makeCluster(n_cores)
  on.exit(stopCluster(cl), add = TRUE)
  
  clusterExport(
    cl,
    varlist = c("iris_data", "coef_function"),
    envir = environment()
  )
  
  clusterSetRNGStream(cl, iseed = 123)
  
  parLapply(
    cl,
    X = seq_len(n_repetitions),
    fun = coef_function
  )
}

# Benchmark
cores_benchmark <- bench::mark(
  core_02 = run_parallel(2),
  core_04 = run_parallel(4),
  core_06 = run_parallel(6),
  core_08 = run_parallel(8),
  core_10 = run_parallel(10),
  core_12 = run_parallel(12),
  core_14 = run_parallel(14),
  core_16 = run_parallel(16),
  iterations = 3,
  check = FALSE,
  memory = FALSE
)

plot(cores_benchmark)

```

# Exercise 4
Using the optimal number of cores in your case, compare (benchmark) the performance of the method you used to solve the previous exercise and the same using mirai and in_parallel instead. For this, modify the script in the previous exercise to add this comparision at the end.

```{r}
# Benchmark with mirai and do_parallel
# libraries
library(mirai)
library(parallel)
library(foreach)
library(doParallel)
library(dplyr)
library(purrr)
library(bench)

# Optimal number of cores according to previous benchmark
optimal_cores <- 8

# Method 1: parallel::parLapply with optimal number of cores

run_parLapply_optimal <- function() {
  
  cl <- parallel::makeCluster(optimal_cores)
  on.exit(parallel::stopCluster(cl), add = TRUE)
  
  parallel::clusterExport(
    cl,
    varlist = c("iris_data", "coef_function", "n_repetitions"),
    envir = environment()
  )
  
  parallel::clusterSetRNGStream(cl, iseed = 123)
  
  parallel::parLapply(
    cl,
    X = seq_len(n_repetitions),
    fun = coef_function
  )
}

# Method 2: purrr::map with mirai::in_parallel

run_mirai_optimal <- function() {
  
  mirai::daemons(optimal_cores)
  on.exit(mirai::daemons(0), add = TRUE)
  
  purrr::map(
    seq_len(n_repetitions),
    purrr::in_parallel(
      \(repetition) {
        
        sample_individuals <- sample(
          seq_len(nrow(iris_data)),
          nrow(iris_data),
          replace = TRUE
        )
        
        boot_data <- iris_data[sample_individuals, ]
        
        model_res <- suppressWarnings(
          stats::glm(
            Species ~ Petal.Length,
            data = boot_data,
            family = stats::binomial()
          )
        )
        
        stats::coefficients(model_res)
      },
      iris_data = iris_data
    )
  )
}

# Benchmark comparison

parallel_comparison_benchmark <- bench::mark(
  parLapply_optimal = run_parLapply_optimal(),
  mirai_in_parallel = run_mirai_optimal(),
  iterations = 3,
  check = FALSE,
  memory = FALSE,
  time_unit = "s"
)

parallel_comparison_benchmark <- data.frame(parallel_comparison_benchmark)

```
* Is there any differences? Answer in a comment in the script.
Yes, results from the benchmark show a improvement in the total time of 5 seconds using mirai over doParallel with 8 cores.

# Ejercicio 5: Parallelization II

### 1) Check the strategy you wrote. Would you change something now after knowing more about loops and parallelization?
Si, en un principio había incluido los pasos de parquet y duckdb para poder leer y cargar los archivos. Ahora incluiría otro paso para paralelizar el proceso mediante mirai y purrr::map

### 2) Write an script that perform the requested task, using purrr::map

```{r}
library(DBI)
library(duckdb)
library(dplyr)
library(purrr)
library(glue)
library(sf)
library(mapSpain)
library(arrow)
library(stringr)
library(tibble)
library(readr)

# Create folders
dir.create("DATA", showWarnings = FALSE)
dir.create("outputs", showWarnings = FALSE)

# 1. Install DuckDB extensions

duckdb_conn <- DBI::dbConnect(duckdb::duckdb())

DBI::dbExecute(duckdb_conn, "INSTALL spatial;")
DBI::dbExecute(duckdb_conn, "LOAD spatial;")
DBI::dbExecute(duckdb_conn, "INSTALL httpfs;")
DBI::dbExecute(duckdb_conn, "LOAD httpfs;")

DBI::dbDisconnect(duckdb_conn, shutdown = TRUE)

# 2. Create daily file URLs for 2020

dates_2020 <- seq.Date(
  from = as.Date("2020-01-01"),
  to   = as.Date("2020-12-31"),
  by   = "day"
)

date_ids <- format(dates_2020, "%Y%m%d")

file_url <- glue::glue(
  "https://data-emf.creaf.cat/public/parquet/stations_data_historical/meteo_stations_2000_2024.parquet"
)

# 3. Prepare municipality geometries
municipalities_sf <- mapSpain::esp_get_munic() |>
  sf::st_transform(crs = 25830)

# Create municipality code
if ("cod_muni" %in% names(municipalities_sf)) {
  municipalities_sf$municipality_code <- municipalities_sf$cod_muni
} else if (all(c("cpro", "cmun") %in% names(municipalities_sf))) {
  municipalities_sf$municipality_code <- paste0(
    municipalities_sf$cpro,
    municipalities_sf$cmun
  )
} else {
  municipalities_sf$municipality_code <- as.character(seq_len(nrow(municipalities_sf)))
}

# Create municipality name
if ("name" %in% names(municipalities_sf)) {
  municipalities_sf$municipality_name <- municipalities_sf$name
} else {
  municipalities_sf$municipality_name <- municipalities_sf$municipality_code
}

municipalities_tbl <- municipalities_sf %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(
    municipality_code,
    municipality_name
  )

municipalities_tbl$wkt <- st_as_text(st_geometry(municipalities_sf))

# Save municipalities as a small local parquet file
municipality_file <- "DATA/municipalities_spain_25830.parquet"

arrow::write_parquet(
  municipalities_tbl,
  municipality_file
)

municipality_file_sql <- normalizePath(
  municipality_file,
  winslash = "/",
  mustWork = TRUE
)

# 4. Inspect metadata
duckdb_conn <- DBI::dbConnect(duckdb::duckdb())

DBI::dbExecute(duckdb_conn, "LOAD httpfs;")

schema <- DBI::dbGetQuery(
  duckdb_conn,
  glue::glue("
    DESCRIBE SELECT *
    FROM read_parquet('{file_url}')
  ")
)

DBI::dbDisconnect(duckdb_conn, shutdown = TRUE)

# 5. Define summary expressions
all_cols <- schema$column_name

summary_candidates <- tibble::tibble(
  column = c(
    "MinTemperature",
    "MaxTemperature",
    "MeanTemperature",
    "MinRelativeHumidity",
    "MaxRelativeHumidity",
    "MeanRelativeHumidity",
    "Precipitation"
  ),
  expression = c(
    "MIN(m.MinTemperature) AS min_temperature",
    "MAX(m.MaxTemperature) AS max_temperature",
    "AVG(m.MeanTemperature) AS mean_temperature",
    "MIN(m.MinRelativeHumidity) AS min_relative_humidity",
    "MAX(m.MaxRelativeHumidity) AS max_relative_humidity",
    "AVG(m.MeanRelativeHumidity) AS mean_relative_humidity",
    "AVG(m.Precipitation) AS mean_precipitation"
  )
)

summary_expressions <- summary_candidates %>% 
  dplyr::filter(column %in% all_cols) %>% 
  dplyr::pull(expression) %>% 
  paste(collapse = ",\n      ")


# 6. Function to process one day

process_day <- function(url, municipality_file_sql, summary_expressions) {
  
  duckdb_conn <- DBI::dbConnect(duckdb::duckdb())
  on.exit(DBI::dbDisconnect(duckdb_conn, shutdown = TRUE), add = TRUE)
  
  DBI::dbExecute(duckdb_conn, "LOAD spatial;")
  DBI::dbExecute(duckdb_conn, "LOAD httpfs;")
  
  query <- glue::glue("
    WITH municipalities AS (
      SELECT *
      FROM read_parquet('{municipality_file_sql}')
    )
    
    SELECT
      mu.municipality_code,
      mu.municipality_name,
      m.dates,
      {summary_expressions},
      COUNT(*) AS n_cells
    FROM read_parquet('{url}') AS m
    INNER JOIN municipalities AS mu
      ON ST_Intersects(m.geom, ST_GeomFromText(mu.wkt))
    GROUP BY
      mu.municipality_code,
      mu.municipality_name,
      m.dates
    ORDER BY
      mu.municipality_code,
      m.dates
  ")
  
  DBI::dbGetQuery(duckdb_conn, query)
}

# 7. Process all days using purrr::map()

time_map <- system.time({
  
  meteo_municipal_list <- purrr::map(
    file_url,
    process_day,
    municipality_file_sql = municipality_file_sql,
    summary_expressions = summary_expressions
  )
  
  meteo_municipal_2020 <- dplyr::bind_rows(meteo_municipal_list)
})

```

### 3) Modify the script to perform the task in parallel. Has it any benefit of time?
```{r}
library(mirai)

# 1. Start mirai daemons
mirai::daemons(6)

withr::defer(mirai::daemons(0))

# 2. Process all days in parallel
time_parallel <- system.time({
  
  meteo_municipal_parallel_list <- purrr::map(
    file_url,
    purrr::in_parallel(
      \(url) {
        process_day(
          url = url,
          municipality_file_sql = municipality_file_sql,
          summary_expressions = summary_expressions
        )
      },
      process_day = process_day,
      municipality_file_sql = municipality_file_sql,
      summary_expressions = summary_expressions
    )
  )
  meteo_municipal_2020_parallel <- dplyr::bind_rows(
    meteo_municipal_parallel_list
  )
})

```

Comparando los dos procesos, el proceso con purr::map reporta un tiempo de carga de 8.67 segundos mientras que el proceso de paralelización muestra un tiempo de carga de 8.22 segundos. Existe cierta mejora en la eficiencia al paralelizar el proceso, aunque en este caso no parece que la diferencia sea muy amplia.


# Ejercicio 6: Real life problem
As for the last exercise, it will be a real-life problem. For solving this, you should use both what you’ve learned in this module and in the previous modules of the whole course and create an script in your git project to perform the following:

  *a) With the meteorological stations data (https://data-emf.creaf.cat/public/parquet/stations_data_historical/) and using what you have learned       in the previous modules, interpolate the precipitation data to continuous for Asturias province for each day of 2015.
  
  *b) For this you should know how to read and extract the data you need from the parquet file, create a grid of 500m for Asturias, and
      interpolate the precipitation data using the method you prefer.
  
```{r}
# Instalar multiples librerias (Si no lo estan)
check.packages <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
packages<-c("openxlsx","sf","glue", "dplyr", "plyr","data.table","stringr",
            "xts","ggpubr","rstudioapi", "lubridate","DescTools","matrixStats",
            "splitstackshape", "mgsub", "textshape", "psych","fuzzyjoin", "irr", 
            "bench", "DBI", "duckdb", "arrow", "geoarrow")
check.packages(packages)


### IMPORTAR LIBRERIAS ###
library(openxlsx)
library(tidyr)
library(zoo)
library(dplyr)
library(data.table)
library(xts)
library(ggpubr)
library(purrr)
library(rstudioapi)
library(lubridate)
library(DescTools)
library(matrixStats)
library(stringr)
library(splitstackshape)
library(purrr)
library(mgsub)
library(irr)
library(janitor)
library(reshape2)
library(tibble)
library(ggplot2)
library(patchwork)
library(glue)
library(sf)
library(bench)
library(DBI)
library(duckdb)
library(arrow)
library(geoarrow)
library(mapSpain)
library(gstat)
library(stars)

# Select directory path
current_folder<-dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(current_folder)

# 1 - GRID
# Extract Asturias boundaries
muni_asturias <- mapSpain::esp_get_prov("Asturias") %>%
  st_transform(25830)

# Create 500 m grid
grid_500m <- st_make_grid(
  muni_asturias,
  cellsize = 500,
  square = TRUE,
  what = "polygons"
) %>%
  st_sf(
    grid_id = seq_along(.),
    geometry = .
  )

# Intersect Asturias boundary to Asturias grid
grid_500m_asturias <- st_intersection(
  grid_500m,
  st_union(muni_asturias)
)

# Create Asturias centroids
grid_points <- st_centroid(grid_500m_asturias)

# 2 - INTERPOLATE DATA

### 1) Open meteo dataset with the selected parameters
data_parquet <- open_dataset("DATA/meteo_stations_2000_2024.parquet")

meteo_parquet <- data_parquet %>% 
  select(day, year, month, station_province, Precipitation, geom) %>%
  filter(year == 2015,
         station_province == "ASTURIAS") %>% 
  # Extract coordinates
  dplyr::collect() %>%
  dplyr::mutate(
    date = as.Date(
      paste(day, month, year, sep = "/"),
      format = "%d/%m/%Y")

# Create coordinates format from parquet data
geom_sfc <- sf::st_as_sfc(structure(meteo_parquet$geom, class = "WKB"), crs = 4326)

meteo_parquet <- st_sf(meteo_parquet %>%
    dplyr::select(
      date, station_province, Precipitation),
  geometry = geom_sfc) %>%
  sf::st_transform(25830)


### Create a column with dates
meteo_parquet$date <- paste0(meteo_parquet$day,"/",meteo_parquet$month,"/",meteo_parquet$year)
meteo_parquet$date <- as.Date(meteo_parquet$date, format = "%d/%m/%Y")

# Sort by dates
meteo_parquet <- meteo_parquet[order(meteo_parquet$date),]
meteo_parquet <- meteo_parquet %>% 
  select(-c(day, month, year)) %>% 
  relocate(date, .before = station_province)


# 2) Visualize the data to detect possible outliers
meteo_plot <- as.data.frame(meteo_parquet)

ggplot(meteo_plot, aes(x = date, y = Precipitation, color = station_province, group = station_province)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c(
  "ASTURIAS" = "lightblue")) +
  scale_x_date(date_breaks = "1 day", date_labels = "%Y-%m-%d", expand = c(0,0)) +
  labs(title = "",
       x = "",
       Y = "Valores") +
  
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey50", linewidth = 0.4),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    axis.text.x = element_text(angle = 45),
    legend.title = element_blank(),
  )


# 3) Calculate descriptive stats before interpolation to filter the outliers
# 3.1) Calculate IQR with positive values higher than 0 (to avoid bias in the upper limits)

meteo_parquet_positive <- meteo_parquet %>% 
  filter(Precipitation >= 0)

q1 <- quantile(meteo_parquet_positive$Precipitation, 0.25, na.rm = TRUE)
q3 <- quantile(meteo_parquet_positive$Precipitation, 0.75, na.rm = TRUE) 
iqr <- q3 - q1

# 3.2) Calculate lower and upper limits
lower_limits <- q1 - 1.5 * iqr
if (lower_limits < 0){
  lower_limits <- 0
}
upper_limits <- q3 + 1.5 * iqr
if (upper_limits < 30){
  upper_limits <- 50
}

# 3.3) Filter possible outliers
meteo_outliers <- meteo_parquet %>% 
  filter(Precipitation > 50)

# 3.4) Filter all precipitation values according to the limits
meteo_parquet_positive <- meteo_parquet_positive %>% 
  mutate(Precipitation = ifelse(Precipitation < lower_limits | Precipitation > upper_limits, NA, Precipitation))


# 4) Representation after filtering
ggplot(data_cleaned, aes(x = date, y = Precipitation, color = station_province, group = station_province)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c(
  "ASTURIAS" = "lightblue")) +
  scale_x_date(date_breaks = "1 day", date_labels = "%Y-%m-%d", expand = c(0,0)) +
  labs(title = "",
       x = "",
       Y = "Valores") +
  
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey50", linewidth = 0.4),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    axis.text.x = element_text(angle = 45),
    legend.title = element_blank(),
  )

# 5. Interpolation to one day using IDW
interpolate_day_idw <- function(selected_date) {
  meteo_data <- meteo_parquet_positive %>%
    filter(
      date == selected_date,
      !is.na(Precipitation),
      Precipitation >= 0
    )
  
  # IDW 
  idw_result <- gstat::idw(
    formula = Precipitation ~ 1,
    locations = meteo_data,
    newdata = grid_points,
    idp = 2
  )
  
  # Join interpolate data to grid
  result_grid <- grid_500m_asturias
  result_grid$date <- selected_date
  result_grid$precipitation_idw <- idw_result$var1.pred
  
  return(result_grid)
}

# 6) Apply interpolation to 2015 daily dates
dates_2015 <- sort(unique(meteo_parquet_positive$date))

meteo_idw_list_2015 <- purrr::map(
  dates_2015,
  interpolate_day_idw
)
  
meteo_idw_2015_grid <- meteo_idw_list_2015 %>%
  purrr::compact() %>%
  dplyr::bind_rows()

````



