---
title: "Actividad 4.2. Big data"
author: "Ismael Lare David"
date: "`r Sys.Date()`"
---

# Ejercicio 3: Parallelization I

### Using the method you prefer, repeat the previous example (coefficient bootstrapping), changing the number of cores, from 2 to the number in your system (parallel::detectCores()).
*Are always the times better as we increase the number of cores?
No, the plot show a point where the higher number of cores does not necessarily correlate with a higher speed. There is a point where the preparation of the data is heavier than the process itself.
*Can your memory hold all examples?
Yes, the process runs successfully for all the configurations.

```{r}
# libraries
library(mirai)
library(parallel)
library(foreach)
library(doParallel)
library(dplyr)

# data needed
n_repetitions <- 1e4
res_coefs <- list()
iris_data <- iris |>
  dplyr::filter(Species != "setosa")
# set the number of cores to use
doParallel::registerDoParallel(cores = 8)
# foreach loop (for illustration purposes, we check the time used)
eight_cores <- system.time(
  {res_coefs <- foreach::foreach(index = 1:n_repetitions) %dopar% {
    sample_individuals <- sample(85, 85, replace = TRUE)
    model_res <- glm(
      iris_data[sample_individuals, "Species"] ~ iris_data[sample_individuals, "Petal.Length"],
      family = binomial
    )
    coefficients(model_res)
  }}
)

test_efficiency <- rbind(four_cores, six_cores, eight_cores)

# Mclapply

# Data
iris_data <- iris |>
  filter(Species != "setosa") |>
  mutate(Species = droplevels(Species))

# Function to process one repetition
coef_function <- function(repetition) {
  
  sample_individuals <- sample(
    seq_len(nrow(iris_data)),
    nrow(iris_data),
    replace = TRUE
  )
  
  boot_data <- iris_data[sample_individuals, ]
  
  model_res <- suppressWarnings(
    glm(
      Species ~ Petal.Length,
      data = boot_data,
      family = binomial
    )
  )
  
  coefficients(model_res)
}

# Number of repetitions
n_repetitions <- 1e4

# Function to run parallel processing with a given number of cores
run_parallel <- function(n_cores) {
  
  cl <- makeCluster(n_cores)
  on.exit(stopCluster(cl), add = TRUE)
  
  clusterExport(
    cl,
    varlist = c("iris_data", "coef_function"),
    envir = environment()
  )
  
  clusterSetRNGStream(cl, iseed = 123)
  
  parLapply(
    cl,
    X = seq_len(n_repetitions),
    fun = coef_function
  )
}

# Benchmark
cores_benchmark <- bench::mark(
  core_02 = run_parallel(2),
  core_04 = run_parallel(4),
  core_06 = run_parallel(6),
  core_08 = run_parallel(8),
  core_10 = run_parallel(10),
  core_12 = run_parallel(12),
  core_14 = run_parallel(14),
  core_16 = run_parallel(16),
  iterations = 3,
  check = FALSE,
  memory = FALSE
)

plot(cores_benchmark)

```

# Exercise 4
Using the optimal number of cores in your case, compare (benchmark) the performance of the method you used to solve the previous exercise and the same using mirai and in_parallel instead. For this, modify the script in the previous exercise to add this comparision at the end.

```{r}
# Benchmark with mirai and do_parallel
# libraries
library(mirai)
library(parallel)
library(foreach)
library(doParallel)
library(dplyr)
library(purrr)
library(bench)

# Optimal number of cores according to previous benchmark
optimal_cores <- 8

# Method 1: parallel::parLapply with optimal number of cores

run_parLapply_optimal <- function() {
  
  cl <- parallel::makeCluster(optimal_cores)
  on.exit(parallel::stopCluster(cl), add = TRUE)
  
  parallel::clusterExport(
    cl,
    varlist = c("iris_data", "coef_function", "n_repetitions"),
    envir = environment()
  )
  
  parallel::clusterSetRNGStream(cl, iseed = 123)
  
  parallel::parLapply(
    cl,
    X = seq_len(n_repetitions),
    fun = coef_function
  )
}

# Method 2: purrr::map with mirai::in_parallel

run_mirai_optimal <- function() {
  
  mirai::daemons(optimal_cores)
  on.exit(mirai::daemons(0), add = TRUE)
  
  purrr::map(
    seq_len(n_repetitions),
    purrr::in_parallel(
      \(repetition) {
        
        sample_individuals <- sample(
          seq_len(nrow(iris_data)),
          nrow(iris_data),
          replace = TRUE
        )
        
        boot_data <- iris_data[sample_individuals, ]
        
        model_res <- suppressWarnings(
          stats::glm(
            Species ~ Petal.Length,
            data = boot_data,
            family = stats::binomial()
          )
        )
        
        stats::coefficients(model_res)
      },
      iris_data = iris_data
    )
  )
}

# Benchmark comparison

parallel_comparison_benchmark <- bench::mark(
  parLapply_optimal = run_parLapply_optimal(),
  mirai_in_parallel = run_mirai_optimal(),
  iterations = 3,
  check = FALSE,
  memory = FALSE,
  time_unit = "s"
)

parallel_comparison_benchmark <- data.frame(parallel_comparison_benchmark)

```
* Is there any differences? Answer in a comment in the script.
Yes, results from the benchmark show a improvement in the total time of 5 seconds using mirai over doParallel with 8 cores.

# Ejercicio 5: Parallelization II

### 1) Check the strategy you wrote. Would you change something now after knowing more about loops and parallelization?
Si, en un principio había incluido los pasos de parquet y duckdb para poder leer y cargar los archivos. Ahora incluiría otro paso para paralelizar el proceso mediante mirai y purrr::map

### 2) Write an script that perform the requested task, using purrr::map

```{r}
library(DBI)
library(duckdb)
library(dplyr)
library(purrr)
library(glue)
library(sf)
library(mapSpain)
library(arrow)
library(stringr)
library(tibble)
library(readr)

# Create folders
dir.create("DATA", showWarnings = FALSE)
dir.create("outputs", showWarnings = FALSE)

# 1. Install DuckDB extensions

duckdb_conn <- DBI::dbConnect(duckdb::duckdb())

DBI::dbExecute(duckdb_conn, "INSTALL spatial;")
DBI::dbExecute(duckdb_conn, "LOAD spatial;")
DBI::dbExecute(duckdb_conn, "INSTALL httpfs;")
DBI::dbExecute(duckdb_conn, "LOAD httpfs;")

DBI::dbDisconnect(duckdb_conn, shutdown = TRUE)

# 2. Create daily file URLs for 2020

dates_2020 <- seq.Date(
  from = as.Date("2020-01-01"),
  to   = as.Date("2020-12-31"),
  by   = "day"
)

date_ids <- format(dates_2020, "%Y%m%d")

file_url <- glue::glue(
  "https://data-emf.creaf.cat/public/parquet/stations_data_historical/meteo_stations_2000_2024.parquet"
)

# 3. Prepare municipality geometries
municipalities_sf <- mapSpain::esp_get_munic() |>
  sf::st_transform(crs = 25830)

# Create municipality code
if ("cod_muni" %in% names(municipalities_sf)) {
  municipalities_sf$municipality_code <- municipalities_sf$cod_muni
} else if (all(c("cpro", "cmun") %in% names(municipalities_sf))) {
  municipalities_sf$municipality_code <- paste0(
    municipalities_sf$cpro,
    municipalities_sf$cmun
  )
} else {
  municipalities_sf$municipality_code <- as.character(seq_len(nrow(municipalities_sf)))
}

# Create municipality name
if ("name" %in% names(municipalities_sf)) {
  municipalities_sf$municipality_name <- municipalities_sf$name
} else {
  municipalities_sf$municipality_name <- municipalities_sf$municipality_code
}

municipalities_tbl <- municipalities_sf %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(
    municipality_code,
    municipality_name
  )

municipalities_tbl$wkt <- st_as_text(st_geometry(municipalities_sf))

# Save municipalities as a small local parquet file
municipality_file <- "DATA/municipalities_spain_25830.parquet"

arrow::write_parquet(
  municipalities_tbl,
  municipality_file
)

municipality_file_sql <- normalizePath(
  municipality_file,
  winslash = "/",
  mustWork = TRUE
)

# 4. Inspect metadata
duckdb_conn <- DBI::dbConnect(duckdb::duckdb())

DBI::dbExecute(duckdb_conn, "LOAD httpfs;")

schema <- DBI::dbGetQuery(
  duckdb_conn,
  glue::glue("
    DESCRIBE SELECT *
    FROM read_parquet('{file_url}')
  ")
)

DBI::dbDisconnect(duckdb_conn, shutdown = TRUE)

# 5. Define summary expressions
all_cols <- schema$column_name

summary_candidates <- tibble::tibble(
  column = c(
    "MinTemperature",
    "MaxTemperature",
    "MeanTemperature",
    "MinRelativeHumidity",
    "MaxRelativeHumidity",
    "MeanRelativeHumidity",
    "Precipitation"
  ),
  expression = c(
    "MIN(m.MinTemperature) AS min_temperature",
    "MAX(m.MaxTemperature) AS max_temperature",
    "AVG(m.MeanTemperature) AS mean_temperature",
    "MIN(m.MinRelativeHumidity) AS min_relative_humidity",
    "MAX(m.MaxRelativeHumidity) AS max_relative_humidity",
    "AVG(m.MeanRelativeHumidity) AS mean_relative_humidity",
    "AVG(m.Precipitation) AS mean_precipitation"
  )
)

summary_expressions <- summary_candidates %>% 
  dplyr::filter(column %in% all_cols) %>% 
  dplyr::pull(expression) %>% 
  paste(collapse = ",\n      ")


# 6. Function to process one day

process_day <- function(url, municipality_file_sql, summary_expressions) {
  
  duckdb_conn <- DBI::dbConnect(duckdb::duckdb())
  on.exit(DBI::dbDisconnect(duckdb_conn, shutdown = TRUE), add = TRUE)
  
  DBI::dbExecute(duckdb_conn, "LOAD spatial;")
  DBI::dbExecute(duckdb_conn, "LOAD httpfs;")
  
  query <- glue::glue("
    WITH municipalities AS (
      SELECT *
      FROM read_parquet('{municipality_file_sql}')
    )
    
    SELECT
      mu.municipality_code,
      mu.municipality_name,
      m.dates,
      {summary_expressions},
      COUNT(*) AS n_cells
    FROM read_parquet('{url}') AS m
    INNER JOIN municipalities AS mu
      ON ST_Intersects(m.geom, ST_GeomFromText(mu.wkt))
    GROUP BY
      mu.municipality_code,
      mu.municipality_name,
      m.dates
    ORDER BY
      mu.municipality_code,
      m.dates
  ")
  
  DBI::dbGetQuery(duckdb_conn, query)
}

# 7. Process all days using purrr::map()

time_map <- system.time({
  
  meteo_municipal_list <- purrr::map(
    file_url,
    process_day,
    municipality_file_sql = municipality_file_sql,
    summary_expressions = summary_expressions
  )
  
  meteo_municipal_2020 <- dplyr::bind_rows(meteo_municipal_list)
})

```