---
title: "exercise_6"
format: html
editor: visual
---

# Ejercicio 6: Real life problem

As for the last exercise, it will be a real-life problem. For solving this, you should use both what you’ve learned in this module and in the previous modules of the whole course and create an script in your git project to perform the following:

\*a) With the meteorological stations data (<https://data-emf.creaf.cat/public/parquet/stations_data_historical/)> and using what you have learned in the previous modules, interpolate the precipitation data to continuous for Asturias province for each day of 2007.

\*b) For this you should know how to read and extract the data you need from the parquet file, create a grid of 500m for Asturias, and interpolate the precipitation data using the method you prefer.

```{r}
# Instalar multiples librerias (Si no lo estan)
check.packages <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
packages<-c("openxlsx","sf","glue", "dplyr", "plyr","data.table","stringr",
            "xts","ggpubr","rstudioapi", "lubridate","DescTools","matrixStats",
            "splitstackshape", "mgsub", "textshape", "psych","fuzzyjoin", "irr", 
            "bench", "DBI", "duckdb", "arrow", "geoarrow")
check.packages(packages)


### IMPORTAR LIBRERIAS ###
library(openxlsx)
library(tidyr)
library(zoo)
library(dplyr)
library(data.table)
library(xts)
library(ggpubr)
library(purrr)
library(rstudioapi)
library(lubridate)
library(DescTools)
library(matrixStats)
library(stringr)
library(splitstackshape)
library(purrr)
library(mgsub)
library(irr)
library(janitor)
library(reshape2)
library(tibble)
library(ggplot2)
library(patchwork)
library(glue)
library(sf)
library(bench)
library(DBI)
library(duckdb)
library(arrow)
library(geoarrow)
library(mapSpain)
library(gstat)
library(stars)

# Select directory path
current_folder<-dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(current_folder)

data_folder <- file.path(current_folder, "..", "DATA")

# 1 - GRID
# Extract Asturias boundaries
muni_asturias <- mapSpain::esp_get_prov("Asturias") %>%
  st_transform(25830)

# Create 500 m grid
grid_500m <- st_make_grid(
  muni_asturias,
  cellsize = 500,
  square = TRUE,
  what = "polygons"
) %>%
  st_sf(
    grid_id = seq_along(.),
    geometry = .
  )

# Intersect Asturias boundary to Asturias grid
grid_500m_asturias <- st_intersection(
  grid_500m,
  st_union(muni_asturias)
)

# Create Asturias centroids
grid_points <- st_centroid(grid_500m_asturias)

# 2 - INTERPOLATE DATA

### 1) Open meteo dataset with the selected parameters
data_parquet <- open_dataset(paste0(data_folder, "/", "meteo_stations_2000_2024.parquet"))

meteo_parquet <- data_parquet %>% 
  select(day, year, month, station_province, Precipitation, geom) %>%
  filter(year == 2007,
         station_province == "ASTURIAS") %>% 
  # Extract coordinates
  dplyr::collect() %>%
  dplyr::mutate(
    date = as.Date(
      paste(day, month, year, sep = "/"),
      format = "%d/%m/%Y"))

# Create coordinates format from parquet data
geom_sfc <- sf::st_as_sfc(structure(meteo_parquet$geom, class = "WKB"), crs = 4326)

meteo_parquet <- st_sf(meteo_parquet %>%
    dplyr::select(
      date, station_province, Precipitation),
  geometry = geom_sfc) %>%
  sf::st_transform(25830)

# Sort by dates
meteo_parquet <- meteo_parquet[order(meteo_parquet$date),]

# 2) Visualize the data to detect possible outliers
meteo_plot <- as.data.frame(meteo_parquet)

ggplot(meteo_plot, aes(x = date, y = Precipitation, color = station_province, group = station_province)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c(
  "ASTURIAS" = "lightblue")) +
  scale_x_date(date_breaks = "1 day", date_labels = "%Y-%m-%d", expand = c(0,0)) +
  labs(title = "",
       x = "",
       Y = "Valores") +
  
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey50", linewidth = 0.4),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    axis.text.x = element_text(angle = 45),
    legend.title = element_blank(),
  )


# 3) Calculate descriptive stats before interpolation to filter the outliers
# 3.1) Calculate IQR with positive values higher than 0 (to avoid bias in the upper limits)

meteo_parquet_positive <- meteo_parquet %>% 
  filter(Precipitation >= 0)

q1 <- quantile(meteo_parquet_positive$Precipitation, 0.25, na.rm = TRUE)
q3 <- quantile(meteo_parquet_positive$Precipitation, 0.75, na.rm = TRUE) 
iqr <- q3 - q1

# 3.2) Calculate lower and upper limits
lower_limits <- q1 - 1.5 * iqr
if (lower_limits < 0){
  lower_limits <- 0
}
upper_limits <- q3 + 1.5 * iqr
if (upper_limits < 30){
  upper_limits <- 50
}

# 3.3) Filter possible outliers
meteo_outliers <- meteo_parquet %>% 
  filter(Precipitation > 50)

# 3.4) Filter all precipitation values according to the limits
meteo_parquet_positive <- meteo_parquet_positive %>% 
  mutate(Precipitation = ifelse(Precipitation < lower_limits | Precipitation > upper_limits, NA, Precipitation))


# 4) Representation after filtering
ggplot(data_cleaned, aes(x = date, y = Precipitation, color = station_province, group = station_province)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c(
  "ASTURIAS" = "lightblue")) +
  scale_x_date(date_breaks = "1 day", date_labels = "%Y-%m-%d", expand = c(0,0)) +
  labs(title = "",
       x = "",
       Y = "Valores") +
  
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey50", linewidth = 0.4),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.y = element_text(size = 12, margin = margin(r = 10)),
    axis.text.x = element_text(angle = 45),
    legend.title = element_blank(),
  )

# 5. Interpolation to one day using IDW
interpolate_day_idw <- function(selected_date) {
  meteo_data <- meteo_parquet_positive %>%
    filter(
      date == selected_date,
      !is.na(Precipitation),
      Precipitation >= 0
    )
  
  # IDW 
  idw_result <- gstat::idw(
    formula = Precipitation ~ 1,
    locations = meteo_data,
    newdata = grid_points,
    idp = 2
  )
  
  # Join interpolate data to grid
  result_grid <- grid_500m_asturias
  result_grid$date <- selected_date
  result_grid$precipitation_idw <- idw_result$var1.pred
  
  return(result_grid)
}

# 6) Apply interpolation to 2015 daily dates
dates_2007 <- sort(unique(meteo_parquet_positive$date))

meteo_idw_list_2007 <- purrr::map(
  dates_2007,
  interpolate_day_idw
)
  
meteo_idw_2007_grid <- meteo_idw_list_2007 %>%
  purrr::compact() %>%
  dplyr::bind_rows()
```
  
  *c) Calculate the mean bias of the interpolation (predicted - observed) for all the period (2007)
  
```{r}
# Read gpkg file
library(sf)
library(dplyr)

# Select directory path
current_folder<-dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(current_folder)

data_folder <- file.path(current_folder, "..", "DATA")
output_folder <- file.path(current_folder, "..", "outputs")

meteo_gpkg <- sf::st_read(
  paste0(output_folder, "/", "precipitation_idw_asturias_2007_2.gpkg"
))

### 1) Open meteo dataset with the selected parameters
data_parquet <- open_dataset(paste0(data_folder, "/", "meteo_stations_2000_2024.parquet"))

meteo_parquet <- data_parquet %>% 
  select(day, year, month, station_province, Precipitation, geom) %>%
  filter(year == 2007,
         station_province == "ASTURIAS") %>% 
  # Extract coordinates
  dplyr::collect() %>%
  dplyr::mutate(
    date = as.Date(
      paste(day, month, year, sep = "/"),
      format = "%d/%m/%Y"))

# Create coordinates format from parquet data
geom_sfc <- sf::st_as_sfc(structure(meteo_parquet$geom, class = "WKB"), crs = 4326)

meteo_parquet <- st_sf(meteo_parquet %>%
    dplyr::select(
      date, station_province, Precipitation),
  geometry = geom_sfc) %>%
  sf::st_transform(25830)

# Sort by dates
meteo_parquet <- meteo_parquet[order(meteo_parquet$date),]


# 3) Calculate descriptive stats before interpolation to filter the outliers
# 3.1) Calculate IQR with positive values higher than 0 (to avoid bias in the upper limits)
meteo_parquet_positive <- meteo_parquet %>% 
  filter(Precipitation >= 0)

q1 <- quantile(meteo_parquet_positive$Precipitation, 0.25, na.rm = TRUE)
q3 <- quantile(meteo_parquet_positive$Precipitation, 0.75, na.rm = TRUE) 
iqr <- q3 - q1

# 3.2) Calculate lower and upper limits
lower_limits <- q1 - 1.5 * iqr
if (lower_limits < 0){
  lower_limits <- 0
}
upper_limits <- q3 + 1.5 * iqr
if (upper_limits < 30){
  upper_limits <- 50
}

# 3.4) Filter all precipitation values according to the limits
meteo_parquet_positive <- meteo_parquet_positive %>% 
  mutate(Precipitation = ifelse(Precipitation < lower_limits | Precipitation > upper_limits, NA, Precipitation))

meteo_observed_2007 <- meteo_parquet_positive %>%
  filter(!is.na(Precipitation)) %>%
  select(date,observed = Precipitation)

meteo_predicted_2007 <- meteo_gpkg %>%
  sf::st_transform(sf::st_crs(meteo_observed_2007)) %>%
  select(
    date,
    predicted = precipitation_idw
  )

# 4) Mean bias
calculate_bias <- function(selected_date) {
  
  observed_day <- meteo_observed_2007 %>%
    dplyr::filter(date == selected_date)
  
  predicted_day <- meteo_predicted_2007 %>%
    dplyr::filter(date == selected_date)
  
  if (nrow(observed_day) == 0 | nrow(predicted_day) == 0) {
    return(tibble::tibble())
  }
  
  # Find nearest coordinate from observed
  nearest_cell <- st_nearest_feature(
    observed_day,
    predicted_day
  )
  
  bias <- tibble::tibble(
    date = observed_day$date,
    observed = observed_day$observed,
    predicted = predicted_day$predicted[nearest_cell],
    bias = predicted - observed
  )
  
  return(bias)
}
# 5) Apply bias for 2007
dates_2007 <- sort(unique(meteo_observed_2007$date))

bias_2007 <- purrr::map_df(
  dates_2007,
  calculate_bias
)

# 6) Calculate mean bias for 2007
mean_bias_2007 <- bias_2007 %>%
  dplyr::summarise(
    n_observations = dplyr::n(),
    mean_observed = mean(observed, na.rm = TRUE),
    mean_predicted = mean(predicted, na.rm = TRUE),
    mean_bias = mean(bias, na.rm = TRUE),
  )
print(mean_bias_2007$mean_bias)
```

The mean bias for the interpolated data is near 0, therefore the predicted data does not over or under estimate the interpolated data.

-   

    d)  Using the official interpolated meteo data <https://data-emf.creaf.cat/public/parquet/daily_interpolated_meteo/> extract the data for Asturias stations location for a sample of 15 days in 2007 and calculate the bias of the official interpolation.

```{r}

##########
# Official interpolated data
# Sample 15 days
set.seed(123)
sample_dates <- sort(sample(dates_2007, 15))

sample_dates_id <- format(
  sample_dates,
  "%Y%m%d"
)

# file urls
file_urls <- glue::glue(
  "https://data-emf.creaf.cat/public/parquet/daily_interpolated_meteo/{sample_dates_id}.parquet"
)
# leon province filter
ast_wkt <- mapSpain::esp_get_prov("Asturias") |>
  sf::st_transform(crs = 25830) |>
  dplyr::pull("geometry") |>
  sf::st_as_text()

# function for each file
process_day <- function(url) {
  # DB conn
  duckdb_conn <- DBI::dbConnect(duckdb::duckdb())
  DBI::dbExecute(duckdb_conn, "INSTALL spatial;")
  DBI::dbExecute(duckdb_conn, "LOAD spatial;")
  DBI::dbExecute(duckdb_conn, "INSTALL httpfs;")
  DBI::dbExecute(duckdb_conn, "LOAD httpfs;")
  withr::defer(DBI::dbDisconnect(duckdb_conn))
  # query
  asturias_query <- glue::glue("
    SELECT
      dates,
      Precipitation,
    ST_AsText(
      CASE
        WHEN ST_GeometryType(geom) IN ('POLYGON', 'MULTIPOLYGON')
          THEN ST_Centroid(geom)
        ELSE geom
      END) AS geometry_wkt
    FROM read_parquet('{url}')
    WHERE
      ST_Intersects(
        geom,
        ST_GeomFromText('{ast_wkt}')
      )
      AND Precipitation IS NOT NULL
  ")
  
  # result
  result <- DBI::dbGetQuery(duckdb_conn, asturias_query)
  
  result_sf <- result |>
    sf::st_as_sf(wkt = "geometry_wkt", crs = 25830, remove = TRUE) |>
    dplyr::rename(date = dates,
                  predicted = Precipitation)
  
  return(result_sf)
}

# Paralellization with mirai
mirai::daemons(6)
# close mirai daemons when session ends
withr::defer(mirai::daemons(0))

# process all days
official_predicted_data <- purrr::map(
    file_urls,
    purrr::in_parallel(
      \(url) {process_day(url)},
      process_day = process_day,
      ast_wkt = ast_wkt
    )
  )

official_predicted_data <- official_predicted_data %>% 
  compact() %>%
  bind_rows()

### Observed data
### 1) Open meteo dataset with the selected parameters
data_parquet <- open_dataset(paste0(data_folder, "/", "meteo_stations_2000_2024.parquet"))

meteo_parquet <- data_parquet %>% 
  select(day, year, month, station_province, Precipitation, geom) %>%
  filter(year == 2007,
         station_province == "ASTURIAS") %>% 
  # Extract coordinates
  dplyr::collect() %>%
  dplyr::mutate(
    date = as.Date(
      paste(day, month, year, sep = "/"),
      format = "%d/%m/%Y"))

# Create coordinates format from parquet data
geom_sfc <- sf::st_as_sfc(structure(meteo_parquet$geom, class = "WKB"), crs = 4326)

meteo_parquet <- st_sf(meteo_parquet %>%
    dplyr::select(
      date, station_province, Precipitation),
  geometry = geom_sfc) %>%
  sf::st_transform(25830)

# Sort by dates
meteo_parquet <- meteo_parquet[order(meteo_parquet$date),]


# 3) Calculate descriptive stats before interpolation to filter the outliers
# 3.1) Calculate IQR with positive values higher than 0 (to avoid bias in the upper limits)
meteo_parquet_positive <- meteo_parquet %>% 
  filter(Precipitation >= 0)

q1 <- quantile(meteo_parquet_positive$Precipitation, 0.25, na.rm = TRUE)
q3 <- quantile(meteo_parquet_positive$Precipitation, 0.75, na.rm = TRUE) 
iqr <- q3 - q1

# 3.2) Calculate lower and upper limits
lower_limits <- q1 - 1.5 * iqr
if (lower_limits < 0){
  lower_limits <- 0
}
upper_limits <- q3 + 1.5 * iqr
if (upper_limits < 30){
  upper_limits <- 50
}

# 3.4) Filter all precipitation values according to the limits
meteo_parquet_positive <- meteo_parquet_positive %>% 
  mutate(Precipitation = ifelse(Precipitation < lower_limits | Precipitation > upper_limits, NA, Precipitation))

# Set dataframes for observed and official predicted data
meteo_observed_2007 <- meteo_parquet_positive %>%
  filter(!is.na(Precipitation)) %>%
  select(date, observed = Precipitation)

meteo_official_predicted_2007 <- official_predicted_data %>%
  sf::st_transform(sf::st_crs(official_predicted_data)) %>%
  select(
    date,
    predicted
  )

# 4) Mean bias
calculate_official_bias <- function(selected_date) {
  
  observed_day <- meteo_observed_2007 %>%
    dplyr::filter(date == selected_date)
  
  predicted_day <- meteo_official_predicted_2007 %>%
    dplyr::filter(date == selected_date)
  
  if (nrow(observed_day) == 0 | nrow(predicted_day) == 0) {
    return(tibble::tibble())
  }
  
  # Find nearest coordinate from observed
  nearest_cell <- st_nearest_feature(
    observed_day,
    predicted_day
  )
  
  bias <- tibble::tibble(
    date = observed_day$date,
    observed = observed_day$observed,
    official_predicted = predicted_day$predicted[nearest_cell],
    bias = official_predicted - observed
  )
  
  return(bias)
}

# 5) Apply bias for 2007
official_bias_2007 <- purrr::map_df(
  sample_dates,
  calculate_official_bias
)

# 6) Calculate mean bias for official data
mean_bias_2007 <- official_bias_2007 %>%
  dplyr::summarise(
    n_observations = dplyr::n(),
    mean_observed = mean(observed, na.rm = TRUE),
    mean_official_predicted = mean(official_predicted, na.rm = TRUE),
    mean_bias = mean(bias, na.rm = TRUE),
  )
print(mean_bias_2007)
```

-   

    e)  Which interpolation, yours or the official, has less bias?, answer with a comment at the end of the script.

```{r}
# Calculate bias for interpolated meteo data in a sample of 15 days in April 2007

meteo_bias_2007 <- purrr::map_df(
  sample_dates,
  calculate_bias
)

# 6) Calculate mean bias for 2007
mean_meteo_bias_2007 <- meteo_bias_2007 %>%
  dplyr::summarise(
    n_observations = dplyr::n(),
    mean_observed = mean(observed, na.rm = TRUE),
    mean_predicted = mean(predicted, na.rm = TRUE),
    mean_bias = mean(bias, na.rm = TRUE),
  )
print(mean_meteo_bias_2007$mean_bias)

 # Calculate bias for official interpolated meteo data in a sample of 15 days in April 2007
official_bias_2007 <- purrr::map_df(
  sample_dates,
  calculate_official_bias
)

# 6) Calculate mean bias for official data
mean_official_bias_2007 <- official_bias_2007 %>%
  dplyr::summarise(
    n_observations = dplyr::n(),
    mean_observed = mean(observed, na.rm = TRUE),
    mean_official_predicted = mean(official_predicted, na.rm = TRUE),
    mean_bias = mean(bias, na.rm = TRUE),
  )
print(mean_official_bias_2007$mean_bias)

```

The bias is greater in official interpolated data (0.40) than in the meteo interpolated data (0.000001). Additional metrics such as RMSE and cross-validations would be required to test the predictive accuracy.
