---
title: "exercise_5"
format: html
editor: visual
---

# Ejercicio 5: Parallelization II

### 1) Check the strategy you wrote. Would you change something now after knowing more about loops and parallelization?

Si, en un principio había incluido los pasos de parquet y duckdb para poder leer y cargar los archivos. Ahora incluiría otro paso para paralelizar el proceso mediante mirai y purrr::map

### 2) Write an script that perform the requested task, using purrr::map

```{r}
library(DBI)
library(duckdb)
library(dplyr)
library(purrr)
library(glue)
library(sf)
library(mapSpain)
library(arrow)
library(stringr)
library(tibble)
library(readr)

# Select directory path
current_folder<-dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(current_folder)

data_folder <- file.path(current_folder, "..", "DATA")

# Create folders
dir.create("DATA", showWarnings = FALSE)
dir.create("outputs", showWarnings = FALSE)

# 1. Install DuckDB extensions

duckdb_conn <- DBI::dbConnect(duckdb::duckdb())

DBI::dbExecute(duckdb_conn, "INSTALL spatial;")
DBI::dbExecute(duckdb_conn, "LOAD spatial;")
DBI::dbExecute(duckdb_conn, "INSTALL httpfs;")
DBI::dbExecute(duckdb_conn, "LOAD httpfs;")

DBI::dbDisconnect(duckdb_conn, shutdown = TRUE)

# 2. Create daily file URLs for 2020

dates_2020 <- seq.Date(
  from = as.Date("2020-01-01"),
  to   = as.Date("2020-12-31"),
  by   = "day"
)

date_ids <- format(dates_2020, "%Y%m%d")

file_url <- glue::glue(
  "https://data-emf.creaf.cat/public/parquet/stations_data_historical/meteo_stations_2000_2024.parquet"
)

# 3. Prepare municipality geometries
municipalities_sf <- mapSpain::esp_get_munic() |>
  sf::st_transform(crs = 25830)

# Create municipality code
if ("cod_muni" %in% names(municipalities_sf)) {
  municipalities_sf$municipality_code <- municipalities_sf$cod_muni
} else if (all(c("cpro", "cmun") %in% names(municipalities_sf))) {
  municipalities_sf$municipality_code <- paste0(
    municipalities_sf$cpro,
    municipalities_sf$cmun
  )
} else {
  municipalities_sf$municipality_code <- as.character(seq_len(nrow(municipalities_sf)))
}

# Create municipality name
if ("name" %in% names(municipalities_sf)) {
  municipalities_sf$municipality_name <- municipalities_sf$name
} else {
  municipalities_sf$municipality_name <- municipalities_sf$municipality_code
}

municipalities_tbl <- municipalities_sf %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(
    municipality_code,
    municipality_name
  )

municipalities_tbl$wkt <- st_as_text(st_geometry(municipalities_sf))

# Save municipalities as a small local parquet file
municipality_file <- paste0(data_folder, "/", "municipalities_spain_25830.parquet")

arrow::write_parquet(
  municipalities_tbl,
  municipality_file
)

municipality_file_sql <- normalizePath(
  municipality_file,
  winslash = "/",
  mustWork = TRUE
)

# 4. Inspect metadata
duckdb_conn <- DBI::dbConnect(duckdb::duckdb())

DBI::dbExecute(duckdb_conn, "LOAD httpfs;")

schema <- DBI::dbGetQuery(
  duckdb_conn,
  glue::glue("
    DESCRIBE SELECT *
    FROM read_parquet('{file_url}')
  ")
)

DBI::dbDisconnect(duckdb_conn, shutdown = TRUE)

# 5. Define summary expressions
all_cols <- schema$column_name

summary_candidates <- tibble::tibble(
  column = c(
    "MinTemperature",
    "MaxTemperature",
    "MeanTemperature",
    "MinRelativeHumidity",
    "MaxRelativeHumidity",
    "MeanRelativeHumidity",
    "Precipitation"
  ),
  expression = c(
    "MIN(m.MinTemperature) AS min_temperature",
    "MAX(m.MaxTemperature) AS max_temperature",
    "AVG(m.MeanTemperature) AS mean_temperature",
    "MIN(m.MinRelativeHumidity) AS min_relative_humidity",
    "MAX(m.MaxRelativeHumidity) AS max_relative_humidity",
    "AVG(m.MeanRelativeHumidity) AS mean_relative_humidity",
    "AVG(m.Precipitation) AS mean_precipitation"
  )
)

summary_expressions <- summary_candidates %>% 
  dplyr::filter(column %in% all_cols) %>% 
  dplyr::pull(expression) %>% 
  paste(collapse = ",\n      ")


# 6. Function to process one day

process_day <- function(url, municipality_file_sql, summary_expressions) {
  
  duckdb_conn <- DBI::dbConnect(duckdb::duckdb())
  on.exit(DBI::dbDisconnect(duckdb_conn, shutdown = TRUE), add = TRUE)
  
  DBI::dbExecute(duckdb_conn, "LOAD spatial;")
  DBI::dbExecute(duckdb_conn, "LOAD httpfs;")
  
  query <- glue::glue("
    WITH municipalities AS (
      SELECT *
      FROM read_parquet('{municipality_file_sql}')
    )
    
    SELECT
      mu.municipality_code,
      mu.municipality_name,
      m.dates,
      {summary_expressions},
      COUNT(*) AS n_cells
    FROM read_parquet('{url}') AS m
    INNER JOIN municipalities AS mu
      ON ST_Intersects(m.geom, ST_GeomFromText(mu.wkt))
    GROUP BY
      mu.municipality_code,
      mu.municipality_name,
      m.dates
    ORDER BY
      mu.municipality_code,
      m.dates
  ")
  
  DBI::dbGetQuery(duckdb_conn, query)
}

# 7. Process all days using purrr::map()

time_map <- system.time({
  
  meteo_municipal_list <- purrr::map(
    file_url,
    process_day,
    municipality_file_sql = municipality_file_sql,
    summary_expressions = summary_expressions
  )
  
  meteo_municipal_2020 <- dplyr::bind_rows(meteo_municipal_list)
})

```

### 3) Modify the script to perform the task in parallel. Has it any benefit of time?

```{r}
library(mirai)

# 1. Start mirai daemons
mirai::daemons(6)

withr::defer(mirai::daemons(0))

# 2. Process all days in parallel
time_parallel <- system.time({
  
  meteo_municipal_parallel_list <- purrr::map(
    file_url,
    purrr::in_parallel(
      \(url) {
        process_day(
          url = url,
          municipality_file_sql = municipality_file_sql,
          summary_expressions = summary_expressions
        )
      },
      process_day = process_day,
      municipality_file_sql = municipality_file_sql,
      summary_expressions = summary_expressions
    )
  )
  meteo_municipal_2020_parallel <- dplyr::bind_rows(
    meteo_municipal_parallel_list
  )
})

```

Comparando los dos procesos, el proceso con purr::map reporta un tiempo de carga de 8.67 segundos mientras que el proceso de paralelización muestra un tiempo de carga de 8.22 segundos. Existe cierta mejora en la eficiencia al paralelizar el proceso, aunque en este caso no parece que la diferencia sea muy amplia.

------------------------------------------------------------------------
